### 2023-03-02 calculus.tex
Include judgmental forms in hazelnut section

The Hazelnut action calculus introduced by \citet{HazelnutPOPL} gives a structure editing scheme
that employs \emph{expression holes}, denoted $\EEHole$, which represent syntactically incomplete
part of the program. Not only are terms always syntactically well-formed, they are always
\emph{well-typed}, ensuring that semantic services are always available. To accomplish this, the
action calculus inserts on-demand \emph{error holes}, denoted $\hhole{\EMV}{}$, which indicate type
errors.

@@

By applying the marked lambda calculus to Hazelnut, the action calculus may be \emph{wholly decoupled}
from the typing semantics. Under this new system, the action calculus is concerned only with the
manipulation of syntax, and the marking procedure yields statically meaningful terms with error
marks, now error holes, inserted at the appropriate positions as a kind of editor decoration. This
solves the problem outlined above; since marking operates on the entire program after each action,
arbitrary constructions are permitted.

Instead of a wholly untyped action calculus, one may re-adapt the existing Hazelnut calculus with
scope-based marking. Action operate over \emph{marked expressions}, and error marks operate as error
holes. When wrapping $\ECMV$ into $\ECLam{x}{\TUnknown}{\ECMV}$, we may re-mark the body under the
context with $x$ binding type $\TUnknown$:
\[%
  \judgment{
    \ctxSynFixedInto{\extendCtx{\ctx}{x}{\TUnknown}}{\erase{\ECMV}}{\ECMV'}{\TMV'}
  }{
    \ASEConLam{\ctx}{\ECMV}{\TMV}{\ECMV'}{\TMV'}{x}
  }{ASEConLam}
\]%
This system optimizes over the untyped version by performing marking only when necessary. In
particular, whenever binders are introduced around an expression, marking redraws error holes. The
supplemental material contains a complete formal description of both variants, and the untyped
action semantics and related metatheory are mechanized in Agda.
% TODO note that we have not verified the metatheory for the typed system

>>>>

The Hazelnut action calculus introduced by \citet{HazelnutPOPL} gives a structure editing scheme
that employs \emph{expression holes}, denoted $\EEHole$, which represent syntactically incomplete
part of the program. Not only are terms always syntactically well-formed, they are always
\emph{well-typed}, ensuring that semantic services are always available. To accomplish this, the
bidirectional action calculus---given by the synthetic action judgment, written
$\ASEAction{\ctx}{\EMV}{\TMV}{\EMV'}{\TMV'}{\AMV}$, and the analytic action judgment, written
$\AAEAction{\ctx}{\EMV}{\EMV'}{\TMV}{\AMV}$---inserts on-demand \emph{error holes}, denoted
$\hhole{\EMV}{}$, which indicate type errors.

@@

By applying the marked lambda calculus to Hazelnut, the action calculus may be \emph{wholly decoupled}
from the typing semantics. Under this new system, the action calculus, given by the singular action
judgment, written $\AUEAction{\EMV}{\EMV'}{\AMV}$, is concerned only with the manipulation of
syntax, and the marking procedure yields statically meaningful terms with error marks, now error
holes, inserted at the appropriate positions as a kind of editor decoration. This solves the problem
outlined above; since marking operates on the entire program after each action, arbitrary
constructions are permitted.

Alternatively, instead of a wholly untyped action calculus, one may re-adapt the existing Hazelnut
calculus with scope-based marking. Action operate over \emph{marked expressions}, and error marks
operate as error holes. The system is once again bidirectional, writing (note $\ECMV$ instead of
$\EMV$) the judgments $\ASEAction{\ctx}{\ECMV}{\TMV}{\ECMV'}{\TMV'}{\AMV}$ and
$\AAEAction{\ctx}{\ECMV}{\ECMV'}{\TMV}{\AMV}$. Then, when wrapping $\ECMV$ into
$\ECLam{x}{\TUnknown}{\ECMV}$, we may simply re-mark the body under the context with $x$ binding
the unknown type:
\[%
  \judgment{
    \ctxSynFixedInto{\extendCtx{\ctx}{x}{\TUnknown}}{\erase{\ECMV}}{\ECMV'}{\TMV'}
  }{
    \ASEConLam{\ctx}{\ECMV}{\TMV}{\ECMV'}{\TMV'}{x}
  }{ASEConLam}
\]%
This system optimizes over the untyped version by performing marking only when necessary. In
particular, whenever binders are introduced around an expression, marking redraws error holes. The
supplemental material contains a complete formal description of both variants, and the untyped
action semantics and related metatheory are mechanized in Agda.

% TODO note that we have not verified the metatheory for the typed system
